---
title: "CUNY SPS DATA 607 Project 1"
author: "Betsy Rosalen"
date: "February 24, 2018"
output:
  html_document:
    theme: cerulean
    toc: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Assignment

In this project, you’re given a text file with chess tournament results where the information has some structure. Your job is to create an R Markdown file that generates a .CSV file (that could for example be imported into a SQL database) with the following information for all of the players:

Player’s Name, Player’s State, Total Number of Points, Player’s Pre-Rating, and Average Pre Chess Rating of Opponents

For the first player, the information would be:

Gary Hua, ON, 6.0, 1794, 1605

1605 was calculated by using the pre-tournament opponents’ ratings of 1436, 1563, 1600, 1610, 1649, 1663, 1716, and dividing by the total number of games played.

## Import the data.

I used read.table so that I could separate data on the pipe "|" symbol.  I tried using the read_delim function from the readr package, but it gave me errors because of the separator rows. 

```{r eval=FALSE, include=FALSE}
library(readr)
theUrl <- "https://raw.githubusercontent.com/betsyrosalen/DATA_607_Data_Acquisition_and_Management/master/Project1/tournamentinfo.txt"
rawChessData <- read_delim(file = theUrl, skip = 1, delim = "|")
str(rawChessData)
```

```{r}
theUrl <- "https://raw.githubusercontent.com/betsyrosalen/DATA_607_Data_Acquisition_and_Management/master/Project1/tournamentinfo.txt"
rawChessData <- read.table(file = theUrl, header = FALSE, skip = 1, sep = "|", fill = TRUE, stringsAsFactors = FALSE)
str(rawChessData)
```

## Combine rows of data that belong to the same observation

The original file has two rows for each player plus a separator row.  I created two data frames, one for each row of data and then combined them into one consolidated data.frame automatically removing the separator rows in the process. 

#### Data frame with row one data

```{r}
chessData1 <- rawChessData[seq(1 , 195, 3), ]
head(chessData1, 5)
```

#### Data frame with row two data

```{r}
chessData2 <- rawChessData[seq(2 , 195, 3), ]
head(chessData2, 5)
```

#### Combined data.frame with one row per player, no separator rows, and column names added.

```{r}
chessData <- cbind(chessData1[-1,1:10], chessData2[-1,1:10])
names(chessData) <- cbind(chessData1[1,1:10], chessData2[1,1:10])
names(chessData)[c(1, 3:10, 11)] <- c('PairNum', 'TotalPts', paste('Round', 1:7, sep=""), 'State')
rownames(chessData) <- NULL
head(chessData, 5)
```

## Create new columns for USCFID, and both ratings

The original file had multiple variables of data in one column, so I used regular expressions to separate out and create new columns for USCFID, and both ratings.

```{r}
library(stringr)
USCFID <- str_extract_all(chessData[12], "\\d{8}")
chessData$USCFID <- unlist(USCFID)
RtgPre <- str_extract_all(chessData[12], "R: +\\d{3,4}")  
RtgPre <- str_extract_all(RtgPre, "\\d{3,4}")
chessData$RtgPre <- as.integer(unlist(RtgPre))
RtgPost <- str_extract_all(chessData[12], "-> *\\d{3,4}")
RtgPost <- str_extract_all(RtgPost, "\\d{3,4}")
chessData$RtgPost <- as.integer(unlist(RtgPost))
head(chessData[c(2,21:23)], 5)
```

## Calculate Average Pre Chess Rating of Opponents

I had a lot of false starts with this section before I figured out how to do this.  I created a separate matrix with the round columns and used a regular expression to pull out just the opponents ID numbers, then used that in a loop to pull the oppenents ratings and calculate the average and add it as a new column to my original data.frame.  

```{r eval=FALSE, include=FALSE}

#opponents <- for (i in 1:nrow(chessData)) {
#  opponents[i,] <- unlist(str_extract_all(chessData[i,4:10],
#                                           "\\d{1,2}"))
#  }

chessData$opponents <- for (i in 1:nrow(chessData)) {chessData$opponents[i,] <- c(unlist(str_extract_all(chessData[i,4:10], "\\d{1,2}")))}
chessData$opponents
#opponents <- table(opponents)
#opponents
```

```{r eval=FALSE, include=FALSE}
chessData$opponents <- for (i in 1:nrow(chessData)) 
  {n <- c(as.numeric(unlist(str_extract_all(chessData[i,4:10], "\\d{1,2}"))))
  chessData$avgRtg[i,] <- as.integer(mean(chessData$RtgPre[n]))
}
```

```{r eval=FALSE, include=FALSE}
library(iterators)
x = iter(chessData, by = "row", 
  unlist(str_extract_all(chessData[,4:10], "\\d{1,2}")))
x
```

```{r eval=FALSE, include=FALSE}
chessData$avg <- rep(0,64)
for (i in 1:64){
  opponents <- c(as.integer(unlist(str_extract_all(chessData[i,4:10],
                                                   "\\d{1,2}"))))
  chessData$avg[,i] <- mean(chessData$RtgPre[opponents], na.rm = TRUE)
}
```

```{r eval=FALSE, include=FALSE}
chessData[24:31] <- chessData[4:10]
for (i in 1:nrow(chessData)){
  for (j in 24:31){
    str_replace(chessData[i ,j],"\\w*\\s*(\\d{1,})","\\1")
  }
}

chessData[24]
```

```{r eval=FALSE, include=FALSE}
opponents <- t(chessData[4:10])
opponents <- as.integer(str_replace(opponents,"\\w*\\s*(\\d{1,})","\\1"))
opponents <- data.frame(matrix(opponents, nrow=64))
#opponents[is.na(opponents)] <- as.numeric(0)
#opponents
x = as.integer(opponents[1,])
class(x)
x
Average <- mean(chessData$RtgPre[x], na.rm = TRUE)  #Why isn't this taking the mean of only row one?!?!?
Average
```

```{r eval=FALSE, include=FALSE}

#  This works!!!  

opponents <- t(chessData[4:10])
opponents <- as.integer(str_replace(opponents,"\\w*\\s*(\\d{1,})","\\1"))
opponents <- matrix(opponents, nrow=64)
opponents[is.na(opponents)] <- as.numeric(0)
#opponents

chessData$avg <- rep(0,64)
x = as.integer(opponents[1:7])
Average <- mean(chessData$RtgPre[x], na.rm = TRUE)
class(Average)
chessData$avg[1] <- Average
```

```{r warning=FALSE}

opponents <- t(chessData[4:10])
opponents <- as.integer(str_replace(opponents,"\\w*\\s*(\\d{1,})","\\1"))
opponents <- matrix(opponents, nrow=64)

x <- 1
for (i in 1:nrow(chessData)){
  y = as.integer(opponents[x:(x+6)])
  Average <- round(mean(chessData$RtgPre[y], na.rm = TRUE))
  chessData$avgOppRating[i] <- Average
  x <- x+7
}
head(chessData[,c(1,2,24)])
```

## Create new data.frame with just the required columns of data

Player’s Name, Player’s State, Total Number of Points, Player’s Pre-Rating, and Average Pre Chess Rating of Opponents

```{r}
finalChessData <- chessData[,c(2,11,3,22,24)]
head(finalChessData)
```

## Export to CSV!

```{r}
write.csv(finalChessData, file = "chess.csv", row.names = FALSE)
```

## Final Thoughts

I struggled with a few sections of this poject, but learned a lot along the way.  I have a feeling if I went back and tried my original ideas again, probably I could get them to work now after all the troubleshooting I did trying different solutions.  I find that it's hard to troublshoot in R.  Error messages are not always as descritive as I am used to from working with Python.  Often I find that my problems are with earlier steps and data that is of the wrong type, rather than errors in the code that gives me the error.  I need to figure out a better way of finding my mistakes and making sure data is the right type.