---
title: "DATA 607 - Data Science in Context Presentation"
author: "Betsy Rosalen"
date: "May 2, 2018"
output: 
  slidy_presentation:
    #theme: cerulean
    css: ./slides.css
editor_options: 
  chunk_output_type: inline
---

```{r global_options, echo=FALSE}
knitr::opts_chunk$set(echo=FALSE, warning=FALSE, message=FALSE)
```

```{r setup, include=FALSE}
# library(knitr)
library(tidyverse)
library(DT)
#library(kableExtra)
#library(likert)
```

# Data Visualization

<br>

## **What this presentation is about**  

- What is data visualization?
- Why do we need it?
- History of data disualization
- What makes a good visuaization?

# Data Visualization

<br>

## **What this presentation is NOT about**  

- How to make/code data visualizations


# What is data visualization?

## Anything that visually communicates any amount of data

- charts
- graphs
- maps 
- diagrams
- emojis!


# Why do we need data visualization?

- Helps us quickly understand huge amounts of data at a glance
- Helps identify correlations, trends, and patterns that might otherwise go unnoticed.
- Helps draw attention to the main message

# History of data visualization



# What makes a good visuaization?


