---
title: "CUNY SPS DATA 607 Assignment 7"
author: "Betsy Rosalen"
date: "April 1, 2018"
output:
  html_document:
    theme: cerulean
    toc: yes
    #css: ./reports.css
editor_options: 
  chunk_output_type: inline
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(XML)
library(RCurl)
library(xml2)
library(jsonlite)
```

### Assignment – Working with XML and JSON in R

Pick three of your favorite books on one of your favorite subjects. At least one of the books should have more than one author. For each book, include the title, authors, and two or three other attributes that you find interesting.

Take the information that you’ve selected about these three books, and separately create three files which store the book’s information in HTML (using an html table), XML, and JSON formats (e.g. “books.html”, “books.xml”, and “books.json”). To help you better understand the different file structures, I’d prefer that you create each of these files “by hand” unless you’re already very comfortable with the file formats.

Write R code, using your packages of choice, to load the information from each of the three sources into separate R data frames. Are the three data frames identical?

Your deliverable is the three source files and the R code. If you can, package your assignment solution up into an .Rmd file and publish to rpubs.com. [This will also require finding a way to make your three text files accessible from the web].

### Read data from HTML file

```{r}
htmlURL <- "https://raw.githubusercontent.com/betsyrosalen/DATA_607_Data_Acquisition_and_Management/master/Assignment7/books.html"
booksHTML <- getURLContent(htmlURL)
HTMLtable <- readHTMLTable(booksHTML)
HTMLtable
```

```{r eval=FALSE, include=FALSE}
as_list(booksHTML)
```


### Read data from XML file

```{r}
xmlURL <- "https://raw.githubusercontent.com/betsyrosalen/DATA_607_Data_Acquisition_and_Management/master/Assignment7/books.xml"
booksXML <- read_xml(xmlURL)
booksXML
```

```{r}
# get all the <record>s
books <- xml_find_all(booksXML, "//book")

# extract and clean all the columns
vals <- trimws(xml_text(books))

# extract and clean (if needed) the area names
labs <- trimws(xml_attr(books, "label"))

# mine the column names from the two variable descriptions
# this XPath construct lets us grab either the <categ…> or <real…> tags
# and then grabs the 'name' attribute of them
cols <- xml_attr(xml_find_all(books, "//data/variables/*[self::categoricalvariable or self::realvariable]"), "name")

# this converts each set of <record> columns to a data frame
# after first converting each row to numeric and assigning
# names to each column (making it easier to do the matrix to data frame conv)
dat <- do.call(rbind, lapply(strsplit(vals, "\ +"),
                                 function(x) {
                                   data.frame(rbind(setNames(as.numeric(x),cols)))
                                 }))

# then assign the area name column to the data frame
dat$area_name <- labs

head(dat)
```


### Read data from JSON file

```{r}
jsonURL <- "https://raw.githubusercontent.com/betsyrosalen/DATA_607_Data_Acquisition_and_Management/master/Assignment7/books.json"
booksJSON <- fromJSON(jsonURL)
booksJSON
```