---
title: "CUNY SPS DATA 607 Project 2"
author: "Betsy Rosalen"
date: "March 11, 2018"
output:
  html_document:
    theme: cerulean
    toc: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Assignment Instructions

*The goal of this assignment is to give you practice in preparing different datasets for downstream analysis work.*

Your task is to:

1. Choose any three of the 'wide' datasets identified in the Week 5 Discussion items. For each of the three chosen datasets:
    - Create a .CSV file (or optionally, a MySQL database!) that includes all of the information included in the dataset. You're encouraged to use a 'wide' structure similar to how the information appears in the discussion item, so that you can practice tidying and transformations as described below.
    - Read the information from your .CSV file into R, and use tidyr and dplyr as needed to tidy and transform your data. [Most of your grade will be based on this step!]
    - Perform the analysis requested in the discussion item.
    - Your code should be in an R Markdown file, posted to rpubs.com, and should include narrative descriptions of your data cleanup work, analysis, and conclusions.

# Dataset Three - NYPD Crime Data

The third dataset contains [historical New York City crime data](http://www1.nyc.gov/site/nypd/stats/crime-statistics/historical.page) from the nyc.gov NYPD website.  

## Load the raw data into R from csv file

I downloaded the data for the “Seven Major Felony Offenses” as an excel file and then saved that as a csv file for importing into R.  The data set contains totals for each of the seven major felony offenses (Murder & Non Negl. Manslaughter, Rape, Robbery, Felony Assault, Burglary, Grand Larceny, and Grand Larceny of Motor Vehicle) for each year and a grand total line for each year for each precinct.

```{r message=FALSE}
library(RCurl) 
x <- getURL("https://raw.githubusercontent.com/betsyrosalen/DATA_607_Data_Acquisition_and_Management/master/Project2/DS3-1-seven-major-felony-offenses-2000-2017.csv") 
crime <-data.frame(read.csv(text=x, header=FALSE, stringsAsFactors = FALSE))
dim(crime)
head(crime, 5)
```

## Fix header and remove blank rows

First let's remove the two blank rows at the top and make the third row the header row.

```{r}
# make the third row the header row
colnames(crime) <- crime[3,]
# then remove the first 3 rows
crime <- crime[-c(1:3),]
# reset the row indexing
rownames(crime) <- NULL
```

OK, now let's check out the top and bottom of our data just to get a sense of what is there...

```{r}
head(crime, 5)
tail(crime[,1:10], 14)
```

## Remove 'Statistical Notes' from the bottom of the dataset

The bottom of the dataset has some info that really belongs in a separate file or dataframe because it is metadata, not actually part of the dataset.  Looks like our data actually ends on row 624 and our "Statistical Notes" start on Row 628, so let's remove those rows and put the "Statistical Notes" in a separate dataframe for now just so we don't lose that info.

```{r}
# Put the Statistical notes in a separate dataframe 
StatisticalNotes <- crime[-(1:627),1]

# Remove the Statistical Notes and blank rows from the bottom of the crime dataframe
crime <- crime[(1:624),]
```

## Fill in the precinct in rows where it's blank

Every 8 rows refers to the "Seven Major Felony Offenses" plus a total row for each precinct in the dataset. The precinct number was only filled in on the first row of each section.  To tidy our data we need to fill it in on the other seven rows for the precinct.

```{r message=FALSE}
library(tidyverse)
for (x in 2:8){
  crime[seq(x,624,8),1] <- crime[seq(1,624,8),1]
}
glimpse(crime)
sample_n(crime, 5)
```

## Remove total rows and convert our wide data to long

The tidyr gather function makes it easy to make our wide data long so that it is easier to analyze.

```{r}
tidycrime <- crime %>%
  filter(CRIME != 'TOTAL SEVEN MAJOR FELONY OFFENSES') %>%
  gather("Year", "n", 3:20)

sample_n(tidycrime, 10)
```

## Last but not least...

Let's look at the 'Statistical Notes' to see if there's anything in there we need to know before doing our analysis.

```{r}
StatisticalNotes
```

There are definitely some notes in there that we should keep in mind as we do our analysis.  For example, "The 121 pct was created on 7-1-2013 from parts of the 120 and 122 precinct.  Therefore statistics for 120 and 122 precincts following 2012 are not comparable to earlier years."  So if we are looking at specific precincts we need to keep things like this in mind.  If we want to look at the entire date range we would have to aggregate those three precincts in order to get relevant statistics.  

## Exploratory data analysis

First let's see the totals by type of crime...

```{r}
tidycrime %>%
  group_by(CRIME) %>%
  summarise(Totals = sum(n, na.rm=TRUE)) %>%
  arrange(desc(Totals))
```

We can get totals by precinct, but without population numbers it's hard to compare one precinct to another.  Also, the numbers are affected by the 'Statistical Notes' we looked at earlier, so this may not be very useful info...

```{r}
tidycrime %>%
  group_by(PCT) %>%
  summarise(Totals = sum(n, na.rm=TRUE)) %>%
  arrange(desc(Totals))
```

Let's look at just the 120, 121, and 122 precincts together since those numbers can't really be separated and let's add the 123rd precint too since that would encompass all of Staten Island, which is where I live...

```{r}
SICrime <- tidycrime %>%
  filter(PCT %in% c('121', '122', '123', '124')) %>%
  group_by(Year, CRIME) %>%
  summarise(Totals = sum(n, na.rm=TRUE)) %>%
  arrange(desc(Year))

SICrime
```

Now let's graph that to see the trend...

```{r}
ggplot(SICrime, aes(x=Year, y=Totals, group=CRIME)) +
  geom_line(aes(color=CRIME)) +
  theme(axis.text.x = element_text(angle = 90, vjust=0.5, size = 8))
```

Now let's compare that to the rest of the city...

```{r}
NotSICrime <- tidycrime %>%
  filter(!(PCT %in% c('121', '122', '123', '124'))) %>%
  group_by(Year, CRIME) %>%
  summarise(Totals = sum(n, na.rm=TRUE)) %>%
  arrange(desc(Year))

NotSICrime
```

Now let's graph that to see if the trends look similar to those we saw on Statne Island...

```{r}
ggplot(NotSICrime, aes(x=Year, y=Totals, group=CRIME)) +
  geom_line(aes(color=CRIME)) +
  theme(axis.text.x = element_text(angle = 90, vjust=0.5, size = 8))
```


