---
title: "CUNY SPS DATA 607 Project 2"
author: "Betsy Rosalen"
date: "March 11, 2018"
output:
  html_document:
    theme: cerulean
    toc: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Assignment Instructions

*The goal of this assignment is to give you practice in preparing different datasets for downstream analysis work.*

Your task is to:

1. Choose any three of the 'wide' datasets identified in the Week 5 Discussion items. For each of the three chosen datasets:
    - Create a .CSV file (or optionally, a MySQL database!) that includes all of the information included in the dataset. You're encouraged to use a 'wide' structure similar to how the information appears in the discussion item, so that you can practice tidying and transformations as described below.
    - Read the information from your .CSV file into R, and use tidyr and dplyr as needed to tidy and transform your data. [Most of your grade will be based on this step!]
    - Perform the analysis requested in the discussion item.
    - Your code should be in an R Markdown file, posted to rpubs.com, and should include narrative descriptions of your data cleanup work, analysis, and conclusions.

# Dataset Three

## Load the Raw Data into R from CSV File

```{R LoadMovieCSV}
library(RCurl) 
x <- getURL("https://raw.githubusercontent.com/betsyrosalen/DATA_607_Data_Acquisition_and_Management/master/Assignment2/Movie%20Ratings.csv") 
movies <-data.frame(read.csv(text=x, header=TRUE, stringsAsFactors = FALSE))
dim(movies)
head(movies, 3)
```

## Use tidyr to 

```{r}
library(tidyr)

```

