---
title: "CUNY SPS DATA 607 Project 2"
author: "Betsy Rosalen"
date: "March 11, 2018"
output:
  html_document:
    theme: cerulean
    toc: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Assignment Instructions

*The goal of this assignment is to give you practice in preparing different datasets for downstream analysis work.*

Your task is to:

1. Choose any three of the 'wide' datasets identified in the Week 5 Discussion items. For each of the three chosen datasets:
    - Create a .CSV file (or optionally, a MySQL database!) that includes all of the information included in the dataset. You're encouraged to use a 'wide' structure similar to how the information appears in the discussion item, so that you can practice tidying and transformations as described below.
    - Read the information from your .CSV file into R, and use tidyr and dplyr as needed to tidy and transform your data. [Most of your grade will be based on this step!]
    - Perform the analysis requested in the discussion item.
    - Your code should be in an R Markdown file, posted to rpubs.com, and should include narrative descriptions of your data cleanup work, analysis, and conclusions.

# Dataset One - Movie Data

## Load the Raw Data into R from CSV File

I am using some informal data collected from friends on 10 Sci-Fi Movies released in 2017 and how they would rate them on a 1-5 scale, 1 meaning they hated it and 5 meaning they loved it.  A zero indicates that they did not see the movie.  

The movie data was collected using a [Google Form](https://docs.google.com/forms/d/e/1FAIpQLSf_vRXoLGs7P0zYY60BdzvQkqAw5gLJIubSwWC1VnQAOSNezQ/viewform?usp=sf_link) and the responses are stored in a [Google Sheet](https://docs.google.com/spreadsheets/d/1K_a87dxLQR91GwVcqkqZK9kjxcYDl1tbK9EX9WvC7bo/edit#gid=601448852) which I then downloaded and saved as a csv file for import into R for simplicity.  In the future I would like to try linking directly to Google Sheets using a googlesheets package that I recently heard about.

```{R LoadMovieCSV}
library(RCurl) 
x <- getURL("https://raw.githubusercontent.com/betsyrosalen/DATA_607_Data_Acquisition_and_Management/master/Assignment2/Movie%20Ratings.csv") 
movies <-data.frame(read.csv(text=x, header=TRUE))
dim(movies)
head(movies, 3)
```

## Use tidyr to move header row to Movie column

The raw data is a 'wide' datatset with movie titles across the top and reviewer names running down the left side.  To 'tidy' the data the movie titles should be moved to a column so that each combination of a reviewer and movie title is one observation.  The tidyr gather function accomplishes this task.

```{r}
library(tidyr)
tidymovies <- gather(movies, key = 'Movie', value = 'Rating', 3:12)
dplyr::sample_frac(tidymovies, 0.03, replace=FALSE)
```

## Remove Timestamp and Number of Movies Seen columns

The timestamp is just a default of Google Forms that records when the form was submitted, but this info really is not relevant to our movie ratings at all, so I am removing it.  Also the number of movies seen should be a function of the other data rather than listed repeatedly on each row, so I am removing that also.

```{r}
library(dplyr)
# get rid of timestamp and Number.of.Movies.Seen columns...
tidymovies <- tidymovies[,-c(1,3)]
dplyr::sample_frac(tidymovies, 0.03, replace=FALSE)
```

## Change zeros to NA's

My original [Movie Ratings](https://docs.google.com/forms/d/e/1FAIpQLSf_vRXoLGs7P0zYY60BdzvQkqAw5gLJIubSwWC1VnQAOSNezQ/viewform?usp=sf_link) form specified that zero meant the reviewer had not seen the movie, so I want to change all zero's to NA's so that they can be easily excluded from calculations.

```{r}
tidymovies$Rating[tidymovies$Rating == 0] <- NA
dplyr::sample_frac(tidymovies, 0.03, replace=FALSE)
```

## Use dplyr to split first and last names and then remove last names

```{r}
tidymovies <- tidymovies %>%
  separate(Name, c("Reviewer", "Last"), sep=" ") %>%
  select(-Last)
dplyr::sample_frac(tidymovies, 0.03, replace=FALSE)
```

## Clean up movie titles

For some reason googlesheets inserted periods in between every word in the Movie titles, so I want to replace all the periods in the movie names with a space to make them look cleaner.

```{r}
library(stringr)
tidymovies$Movie <- str_replace_all(tidymovies$Movie, "\\.+", " ")
library(DT)
datatable(tidymovies)
```

## A note about NA values

I thought about removing all the NA's from my dataset, but decided not to do that since I had two respondents who didn't see any of the movies and removing all lines with no rating would have removed those two reviewers from my dataset completely.  I thought it was important to note that two of my friends hadn't seen any of the 10 movies.  Obviously they aren't Sci-Fi fans!

## Summary analysis with dplyr and ggplot2

First I want to see the average rating for each movie, but keeping in mind that some movies were only seen by a few of the repondents so their average rating may be biased by the lack of response.  

```{r}
avg_rating_by_movie <- tidymovies %>% 
  group_by(Movie) %>%
  filter(!is.na(Rating)) %>%
  summarise(Avg_Rating = round(mean(Rating, na.rm=TRUE),2), 
            Num_Reviews = n_distinct(Reviewer)) %>%
  arrange(desc(Avg_Rating))
datatable(avg_rating_by_movie)
avg_rating_by_movie %>% arrange(desc(Num_Reviews))
```

The numbers above confirm that the highest rated films are also the films with the highest viewing numbers.  All 7 of the highest rated movies are the same 7 movies with the greatest number of reviewers. My guess is it's likely the movies with low attendance numbers have low attendance because people pre-judge the films based on trailers and other advertisements.  

Maybe viewers already know or believe they know with a high level of certainty whether they will like or not like a film before they even go see it and watching the film only confirms what they already thought anyway in most cases.  If reviewers were not pre-judging films accurately, you wold expect there to be more low ratings where people thought they wouold like a film and then were disappointed.

```{r}
library(ggplot2)
ggplot(avg_rating_by_movie) +
  geom_point(mapping = aes(x=Avg_Rating, y=Movie, color=Num_Reviews))
```

There are few low ratings in the dataset at all.  You can see belwo that therre was only 1 rating of 1 in the entire dataset and 7 ratings of 2 out of 108 total reviews.  I wonder if more reviewers had watched more of the films if there would be more low ratings?

```{r}
tidymovies %>% 
  filter(!is.na(Rating)) %>%
  summarise(n())
tidymovies %>% 
  filter(!is.na(Rating)) %>%
  group_by(Rating) %>%
  summarise(n())
```

The average rating for all movies in the dataset is 3.98

```{r}
tidymovies %>% summarise(round(mean(Rating, na.rm=TRUE),2))
```

```{r}
# Average rating given by each reviewer
avg_rating_by_reviewer <- tidymovies %>% 
  group_by(Reviewer) %>%
  filter(!is.na(Rating)) %>%
  summarise(Avg_Rating = round(mean(Rating, na.rm=TRUE),2),
            Num_Movies = n_distinct(Movie)) %>%
  arrange(desc(Avg_Rating))
avg_rating_by_reviewer
avg_rating_by_reviewer %>% arrange(desc(Num_Movies))
```

```{r}
library(ggplot2)
ggplot(avg_rating_by_reviewer) +
  geom_point(mapping = aes(x=Avg_Rating, y=Reviewer, color=Num_Movies))
```